Reanalysis Step 7: Alternative Propensity Score Specifications

### 2025-08-31 reanalysis-3-imputation_diagnostics.qmd
file naming and logging consistency

cat(sprintf("Dataset: %d eyes from %d patients\n", n_eyes, n_patients))
cat(sprintf("Bilateral patients: %d (%.1f%%)\n", n_bilateral, 100*n_bilateral/n_patients))
cat(sprintf("Missing outcomes: %d eyes (%.1f%%)\n", n_missing_outcome, 100*n_missing_outcome/n_eyes))
cat(sprintf("Imputations performed: %d\n", n_imputations))
```

@@

  print(p_bilateral)
} else {
  cat("Between-eye correlation plot could not be generated.\n")
}
```

@@

cat(sprintf("\nChi-square test: χ² = %.2f, p = %.4f\n", 
            chi_test$statistic, chi_test$p.value))
if(chi_test$p.value < 0.05) {
  cat("The proportions of improved/stable/worsened differ significantly across groups.\n")
} else {
  cat("The proportions are similar across groups.\n")
}
```

@@

```{r summary-assessment}
cat("### Key Findings:\n\n")

# 1. Clustering
cor_preserved <- abs(cor_outcome_imp - cor_baseline) < 0.2
if (cor_preserved) {
  cat(sprintf("✓ Between-eye correlation preserved (%.3f baseline → %.3f imputed)\n", 
              cor_baseline, cor_outcome_imp))
} else {
  cat(sprintf("⚠ Between-eye correlation changed substantially (%.3f → %.3f)\n",
              cor_baseline, cor_outcome_imp))
}

# 2. Distribution similarity
if (ks_test$p.value > 0.05) {
  cat("✓ Imputed and observed distributions are statistically similar\n")
} else {
  cat("⚠ Imputed and observed distributions differ significantly\n")
}

# 3. Clinical relationships
if (all(effect_sizes$diff_outcome > 0)) {
  cat("✓ Lapsed patients have worse outcomes in all cohorts (clinically expected)\n")
} else {
  cat("⚠ Unexpected lapse-outcome relationship in some cohorts\n")
}

# 4. Temporal consistency
if (all(temporal_summary$mean_discrepancy < 0.001)) {
  cat("✓ Temporal arithmetic is preserved\n")
} else {
  cat("⚠ Minor temporal discrepancies detected\n")
}

# 5. MI stability
if (all(mi_variance$lambda < 0.5)) {
  cat("✓ Fraction of missing information is acceptable (<0.5)\n")
} else {
  cat("⚠ High fraction of missing information in some groups\n")
}

@@

@@

>>>>

log_message(sprintf("Dataset: %d eyes from %d patients", n_eyes, n_patients))
log_message(sprintf("Bilateral patients: %d (%.1f%%)", n_bilateral, 100*n_bilateral/n_patients))
log_message(sprintf("Missing outcomes: %d eyes (%.1f%%)", n_missing_outcome, 100*n_missing_outcome/n_eyes))
log_message(sprintf("Imputations performed: %d", n_imputations))
```

@@

  print(p_bilateral)
} else {
  log_message("Between-eye correlation plot could not be generated.")
}
```

@@

log_message(sprintf("Chi-square test: χ² = %.2f, p = %.4f", 
            chi_test$statistic, chi_test$p.value))
if(chi_test$p.value < 0.05) {
  log_message("The proportions of improved/stable/worsened differ significantly across groups.")
} else {
  log_message("The proportions are similar across groups.")
}
```

@@

```{r summary-assessment}
# Key Findings section
cat("### Key Findings:\n\n")

# 1. Clustering
cor_preserved <- abs(cor_outcome_imp - cor_baseline) < 0.2
if (cor_preserved) {
  cat(sprintf("✓ Between-eye correlation preserved (%.3f baseline → %.3f imputed)\n", 
              cor_baseline, cor_outcome_imp))
  log_message(sprintf("Between-eye correlation preserved (%.3f baseline → %.3f imputed)", 
              cor_baseline, cor_outcome_imp))
} else {
  cat(sprintf("⚠ Between-eye correlation changed substantially (%.3f → %.3f)\n",
              cor_baseline, cor_outcome_imp))
  log_message(sprintf("WARNING: Between-eye correlation changed substantially (%.3f → %.3f)",
              cor_baseline, cor_outcome_imp))
}

# 2. Distribution similarity
if (ks_test$p.value > 0.05) {
  cat("✓ Imputed and observed distributions are statistically similar\n")
  log_message("Imputed and observed distributions are statistically similar")
} else {
  cat("⚠ Imputed and observed distributions differ significantly\n")
  log_message("WARNING: Imputed and observed distributions differ significantly")
}

# 3. Clinical relationships
if (all(effect_sizes$diff_outcome > 0)) {
  cat("✓ Lapsed patients have worse outcomes in all cohorts (clinically expected)\n")
  log_message("Lapsed patients have worse outcomes in all cohorts (clinically expected)")
} else {
  cat("⚠ Unexpected lapse-outcome relationship in some cohorts\n")
  log_message("WARNING: Unexpected lapse-outcome relationship in some cohorts")
}

# 4. Temporal consistency
if (all(temporal_summary$mean_discrepancy < 0.001)) {
  cat("✓ Temporal arithmetic is preserved\n")
  log_message("Temporal arithmetic is preserved")
} else {
  cat("⚠ Minor temporal discrepancies detected\n")
  log_message("WARNING: Minor temporal discrepancies detected")
}

# 5. MI stability
if (all(mi_variance$lambda < 0.5)) {
  cat("✓ Fraction of missing information is acceptable (<0.5)\n")
  log_message("Fraction of missing information is acceptable (<0.5)")
} else {
  cat("⚠ High fraction of missing information in some groups\n")
  log_message("WARNING: High fraction of missing information in some groups")
}

@@

log_message(sprintf("Excluded cohort has %.1f%% missing outcomes",
            100 - cohort_summary$has_outcome[cohort_summary$cohort_id == "excluded"] / 
            cohort_summary$n_eyes[cohort_summary$cohort_id == "excluded"] * 100))
log_message(sprintf("Lapse rate: excluded cohort (%.1f%%) vs submitted (%.1f%%)",
            cohort_summary$pct_lapsed[cohort_summary$cohort_id == "excluded"],
            cohort_summary$pct_lapsed[cohort_summary$cohort_id == "submitted"]))

@@

log_message("Recommendations: 1) 2l.pan accounts for clustering, 2) Consider sensitivity analyses, 3) Report by cohort if needed, 4) Use robust SEs")

### 2025-08-31 reanalysis-4-propensity_scores.qmd
test pscore
file naming and logging consistency

---
title: "Reanalysis Step 3: Propensity Score Analysis with Twang"
subtitle: "Calculating IPTW weights using GBM across imputed datasets"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

@@

# Use the reanalysis formula that includes dr_severity
ps_formula <- ps_formula_reanalysis
cat("\nUsing reanalysis formula with dr_severity for propensity score estimation.\n")
```

@@

# Generate diagnostic plots for selected imputations
for (i in imps_to_show) {
  cat("\n### Imputation", i, "\n\n")

@@

  }, error = function(e) {
    # Fallback to simpler summary if bal.table fails
    cat("Note: Detailed balance table unavailable. Showing basic summary:\n\n")

@@

    # Try to extract key metrics from the summary
    if (!is.null(sum_output)) {
      print(sum_output)
    } else {
      cat("Balance statistics not available for this imputation.\n")
    }
  })

  cat("\n\n")
}
```

>>>>

---
title: "Reanalysis Step 4: Propensity Score Analysis with Twang"
subtitle: "Calculating IPTW weights using GBM across imputed datasets"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

@@

# Use the reanalysis formula that includes dr_severity
ps_formula <- ps_formula_reanalysis
log_message("Using reanalysis formula with dr_severity for propensity score estimation.")
```

@@

# Generate diagnostic plots for selected imputations
for (i in imps_to_show) {
  log_message(paste0("Processing diagnostics for Imputation ", i))
  cat("\n### Imputation", i, "\n\n")

@@

  }, error = function(e) {
    # Fallback to simpler summary if bal.table fails
    log_message("Note: Detailed balance table unavailable. Showing basic summary.")

@@

    # Try to extract key metrics from the summary
    if (!is.null(sum_output)) {
      print(sum_output)
    } else {
      log_message("Balance statistics not available for this imputation.")
    }
  })

  # Spacing
}
```

### 2025-08-31 reanalysis-5-outcome_analysis-interaction.qmd
file naming and logging consistency

---
title: "Reanalysis Step 4: Outcome Analysis with Three-Way Interaction"
subtitle: "IPTW-weighted analysis of lapse × DR severity × treatment interactions"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

>>>>

---
title: "Reanalysis Step 5: Outcome Analysis with Three-Way Interaction"
subtitle: "IPTW-weighted analysis of lapse × DR severity × treatment interactions"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

### 2025-08-31 reanalysis-5-outcome_analysis.qmd
file naming and logging consistency

---
title: "Reanalysis Step 4: Main Effects Outcome Analysis"
subtitle: "Analyzing lapse effect without interactions"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
---

>>>>

---
title: "Reanalysis Step 5: Main Effects Outcome Analysis"
subtitle: "Analyzing lapse effect without interactions"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
---

### 2025-08-31 reanalysis-6-sensitivity.qmd
file naming and logging consistency

---
title: "Reanalysis Step 5: Sensitivity Analysis and E-values"
subtitle: "Assessing robustness of main effect findings"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
---

>>>>

---
title: "Reanalysis Step 6: Sensitivity Analysis and E-values"
subtitle: "Assessing robustness of main effect findings"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
---

### 2025-08-31 reanalysis-7-alternative_specifications.qmd
file naming and logging consistency

---
title: "Reanalysis Step 6: Alternative Propensity Score Specifications"
subtitle: "Comprehensive comparison of PS methods with multiple imputation"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 8
---

>>>>

---
title: "Reanalysis Step 7: Alternative Propensity Score Specifications"
subtitle: "Comprehensive comparison of PS methods with multiple imputation"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 8
---
